Reducing retrieval error in MAIAC AOD with the `Earth_obs_cleaning` machine-learning workflow: Results from CONUS

### 2023-08-26 writing/CONUS_AOD.qmd
Label more figures and tweak RMSE map.

@tbl-cv-overall includes the performance metrics for the full cross-validated training data set. Here **raw** is MCD19A2.V061 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.

@@

### Performance by NOAA climate regions

@@

```{r}
# performance by NOAA climate regions
by.region = (dt
    [, eval(performance.j), by = state.noaa.region]
    [order(`Proportion of raw MSE`)])
flextable(pretty.table.numbers(by.region))
```

```{r noaa_region_plot, dev='png', dpi=300}
## map of performance by NOAA regions
conus$state.noaa.region <- factor(
   sapply(conus$STUSPS, USE.NAMES = F,
       function(sa) if (sa %in% c("AK", "HI")) NA else
           names(noaa.regions)[which(sapply(noaa.regions, function(r)
               sa %in% r))]),
   levels = names(noaa.regions))
conus <- merge(conus, by.region, by = "state.noaa.region")
conus_melt <- melt(as.data.table(conus[, c("geometry", "state.noaa.region", "RMSE, raw", "RMSE, corrected")]),
                   id.vars = c("geometry", "state.noaa.region"))
levels(conus_melt$variable) <- c("MAIAC AOD", "corrected AOD")
map_region_rmse <- ggplot(conus_melt, aes(geometry = geometry)) +
  geom_sf(aes(fill = value)) +
  facet_wrap(~variable, ncol = 1) +
  coord_sf(default_crs = st_crs(conus)) +
  scale_fill_continuous("Regional\nRMSE") +
  theme_minimal() +
  theme(axis.text = element_blank(),
        panel.grid.major = element_blank())
map_region_rmse
```

### Performance by month of year

```{r}
result <- dt[,
    keyby = .(Month = month(lubridate::as_datetime(time.sat))),
    eval(performance.j)]
result[, Month := month.abb[Month]]
flextable(pretty.table.numbers(result))
```

@@

```{r agreement_plot}
#| fig-cap: !expr caption
caption = with(list(d = agreement.plot.data(dt)), sprintf(
    "Not shown: %s %s and %s %s",
    d[!in.region & comparison == "y.sat", .N],
    "points outside left panel",
    d[!in.region & comparison == "y.ground.pred", .N],
    "points outside right panel"))
IG(tar_read(agreement.plot.path))
```

>>>>

@tbl-cv-overall includes the performance metrics for the full cross-validated training data set. Here **raw** is Collection 6.1 MAIAC AOD 470nm; **ground** is AERONET AOD interpolated to 470nm; and **corrected** is MAIAC after applying an XGBoost-modeled correction term.

@@

@@

```{r}
#| label: tbl-cv-noaa-regions
#| tbl-cap: Performance by NOAA climate regions. Sorted by decreasing relative improvement in MSE after correction.
by.region = (dt
    [, eval(performance.j), by = state.noaa.region]
    [order(`Proportion of raw MSE`)])
flextable(pretty.table.numbers(by.region))
```

```{r noaa_region_plot, dev='png', dpi=300}
#| label: fig-cv-noaa-regions
#| fig-cap: Map of performance (RMSE) by NOAA regions.
conus$state.noaa.region <- factor(
   sapply(conus$STUSPS, USE.NAMES = F,
       function(sa) if (sa %in% c("AK", "HI")) NA else
           names(noaa.regions)[which(sapply(noaa.regions, function(r)
               sa %in% r))]),
   levels = names(noaa.regions))
conus <- merge(conus, by.region, by = "state.noaa.region")
conus_melt <- melt(as.data.table(conus[, c("geometry", "state.noaa.region", "RMSE, raw", "RMSE, corrected")]),
                   id.vars = c("geometry", "state.noaa.region"))
levels(conus_melt$variable) <- c("MAIAC AOD", "Corrected AOD")
map_region_rmse <- ggplot(conus_melt, aes(geometry = geometry)) +
  geom_sf(aes(fill = value)) +
  facet_wrap(~variable, ncol = 1) +
  coord_sf(default_crs = st_crs(conus)) +
  scale_fill_continuous("Regional\nRMSE", trans = 'reverse') +
  theme_minimal() +
  theme(axis.text = element_blank(),
        panel.grid.major = element_blank(),
        strip.text = element_text(size = 13))
map_region_rmse
```
@fig-cv-noaa-regions shows that performance improves in all regions and leads to greater consistency in the RMSE across CONUS with slightly higher RMSE remaining in the Northwest and West regions.

```{r}
#| label: tbl-cv-month
#| tbl-cap: Performance by month of year.
result <- dt[,
    keyby = .(Month = month(lubridate::as_datetime(time.sat))),
    eval(performance.j)]
result[, Month := month.abb[Month]]
flextable(pretty.table.numbers(result))
```

@@

```{r agreement_plot}
#| label: fig-agreement-plot
#| fig-cap: !expr caption
caption = with(list(d = agreement.plot.data(dt)), sprintf(
    "Not shown: %s %s and %s %s",
    d[!in.region & comparison == "y.sat", .N],
    "points outside left panel",
    d[!in.region & comparison == "y.ground.pred", .N],
    "points outside right panel"))
IG(tar_read(agreement.plot.path))
```
